---
title: "FLR Tutorials - Session 2"
subtitle: "Manipulating, summarizing and operating on FLR objects"
author: Iago MOSQUEIRA (WMR)
date: "[FLR Project](https://flr-project.org)"
output:
  ioslides_presentation:
    widescreen: TRUE
---

<style>
  pre {
    font-size: 15px;
  }
</style>


```{r setup, echo=FALSE, result="hide"}
library(knitr)
library(FLCore)
library(ggplotFL)
data(ple4)
flq <- FLQuant(runif(9), dimnames=list(age=0:2, year=2020:2022), units="kg")
opts_chunk$set(echo=TRUE)
```

## Quick review of FLQuant and FLStock

```{r, echo=TRUE}
FLQuant(runif(9), dimnames=list(age=0:2, year=2020:2022), units="last")
```

## Quick review of FLQuant and FLStock

```{r, echo=FALSE}
plot(ple4) + ggtitle("ple.27.420")
```

## Quick review of FLQuant and FLStock

- `summary(ple4)`
- `catch(ple4)`
- `FLQuant()`
- `ssb(ple4)`
- `dimnames(stock.n(ple4))`
- `units`
- `catch(ple4)[, 1:20]`
- `as.data.frame(ple4)`

## Subsetting

- Like R's `array` but
  - No need to write all the commas
  - By default `drop = FALSE`

```{r}
dim(flq[, 2:3])
dim(flq[, 2:3, drop=TRUE])
```

## Subsetting

- Like R's `array` but
  - No need to write all the commas
  - By default `drop = FALSE`
- By position, `flq[1,]` or name, `flq['0',]`

<div class="columns-2">
```{r}
flq[1, 2:3]
```

```{r}
flq["1", 2:3]
```
</div>

## Subsetting

- Like R's `array` but
  - No need to write all the commas
  - By default `drop = FALSE`
- By position, `flq[1,]` or name, `flq['0',]`
- `window()` for year, also expands

```{r}
window(flq, end=2024)
```

## Subsetting

- Like R's `array` but
  - No need to write all the commas
  - By default `drop = FALSE`
- By position, `flq[1,]` or name, `flq['0',]`
- `window()` for year, also expands
- `expand` works for all dimensions

```{r}
dimnames(expand(flq, age=0:4))$age
```

## Appending

- `append` to join along year
- One `bind` method for each dimension
  - `qbind`
  - `ybind`
  - `ubind`
  - `sbind`
  - `abind`
  - `ibind`

## Arithmetic

- As R's `array` but `iter` need not match (1 vs. N)

```{r}
fla <- FLQuant(runif(90), dim=c(3,3,1,1,1,10))
flb <- FLQuant(runif(9), dim=c(3,3,1,1,1,1))

fla * flb
```

## Summaries along dimensions

- `yearSums`, `yearMeans`, `yearVars`
- `iterMedians`, 'iterCVs'
- Or use `apply(flq, dims, function, ...)`

```{r}
apply(fla, c(2:6), function(x) sum(x <= 0.5))
```

## Dealing with iterations

- `plot()` shows median and quantiles (80%, 33%)

```{r}
flqi <- rnorm(200, catch(ple4), 20000)
plot(flqi)
```

## Dealing with iterations

- `propagate` to move from 1 to N
- `combine` to merge along `iter`
- RNG functions can take FLQuant inputs

```{r}
rnorm(200, flb, 5)
```

## FLStock metrics

- Spawning stock biomass, `ssb()`
- Recruitment, `rec()`
- Total spawning biomass, `tsb()`,
- Vulnerable biomass, `vb()`
- $\bar{F}$, `fbar()`
- $F_{apex}$, `fapex()`
- Surplus production, `sp()`

## Applying functions

- `apply` on FLQuant, as in array
- `qapply` along all FLQuant slots

```{r}
qapply(ple4, function(x)  sum(x == 0))
```

## Applying functions

- `apply` on FLQuant, as in array
- `qapply` along all FLQuant slots
- `lapply` for lists of FLR objects

```{r}
flqs <- FLQuants(A=fla, B=flb)
lapply(flqs, dim)
```

## Applying functions

- `apply` on FLQuant, as in array
- `qapply` along all FLQuant slots
- `lapply` for lists of FLR objects
- `Reduce` to combine list elements

```{r}
Reduce("+", flqs)
```

## Applying functions

- `apply` on FLQuant, as in array
- `qapply` along all FLQuant slots
- `lapply` for lists of FLR objects
- `Reduce` to combine list elements
- `Map` to apply along two or more lists

## FL lists

- Each class has its plural, `FLQuant` -- `FLQuants`
- Any number of elements
- All elements of the same class
- FLQuants, FLStocks, FLIndices

## FLIndices, `ple4.indices`

```{r}
data(ple4.indices)
plot(ple4.indices)
```
