---
title: "FLR Tutorials - Session 6"
subtitle: "Stock assessment methods in FLR"
author: Iago MOSQUEIRA (WMR)
date: "[FLR Project](https://flr-project.org)"
output:
  ioslides_presentation:
    widescreen: TRUE
---

<style>
  pre {
    font-size: 15px;
  }
</style>

```{r setup, echo=FALSE, result="hide"}
library(knitr)
library(ggplotFL)
library(FLXSA)
opts_chunk$set(echo=TRUE)
```

## Stock assessment

<div class="centered">
```{r, echo=FALSE, fig.height=7, fig.width=8, out.width="60%"}
plot(ple4)
```
</div>

## Methods available

- **FLa4a**, Assessment for all. SCA using splines andb submodels.
- **FLXSA**, eXtended Survivor Analysis. VPA + tuning.
- **FLSAM**, Stock Assessment Model. SCA with random walks and mixed effects.
- **AAP**, Aarts & Poos. SCA for sole and plaice, reconstructs discards.
- **BBM**, Two-Stage Stock Assessment Model.
- **mpb**, ASPIC and pella surplus production models in ADMB.

### Other relevant packages
- **ss3om**, Load and work with SS3 outputs.


## Standard interface

### Run

```{r, eval=FALSE}
fit  <- FLXSA(ple4, ple4.indices, FLXSA.control(shk.n=FALSE))
```

### Update (harvest, stock.n)

```{r, eval=FALSE}
run <- ple4 + fit
```

### Diagnostics, residuals

```{r, eval=FALSE}
diagnostics(fit)
residuals(fit)
```

## a4a interface

### Submodels

```{r eval=FALSE}
fmod <- ~ factor(age) + factor(year)
srmod <- ~ factor(year)
```

### Run
```{r}
fit <- sca(stock = ple4, indices = indices, fmodel=fmod, srmodel=srmod)
```

### Update

```{r, eval=FALSE}
run <- ple4 + fit
```

